---
title: 'Example of MultiCOAP'
author: "Wei Liu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example of MultiCOAP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette introduces the usage of MultiCOAP for the analysis of high-dimensional count data with additional high-dimensional covariates, by comparison with other methods. 


The package can be loaded with the command, and define some metric functions:
```{r  eval = TRUE}
library(MultiCOAP)
library(GFM)
normvec <- function(x) sqrt(sum(x^2)/ length(x))

trace_statistic_fun <- function(H, H0){
  
  tr_fun <- function(x) sum(diag(x))
  mat1 <- t(H0) %*% H %*% qr.solve(t(H) %*% H) %*% t(H) %*% H0
  
  tr_fun(mat1) / tr_fun(t(H0) %*% H0)
  
}
trace_list_fun <- function(Hlist, H0list){
  trvec <- rep(NA, length(Hlist))
  for(i in seq_along(trvec)){
    trvec[i] <- trace_statistic_fun(Hlist[[i]], H0list[[i]])
  }
  return(mean(trvec))
}
```

## Generate the simulated data
First, we generate the data simulated data. 
```{r  eval = TRUE}
p <- 100; nvec <- c(200,300); d <- 10
methodNames <- c("MultiCOAP", "COAP", "MSFRlog", "MSFR")
metricMat <- matrix(NA, nrow=length(methodNames), ncol=6)
colnames(metricMat) <- c('A_tr', 'B_tr', 'beta_er', 'F_tr', 'H_tr', 'Time')
row.names(metricMat) <- methodNames

q <- 3; qs <- c(2,2); rank0 <- 2
rho<-c(2,3.5,0.1) 
datList <- gendata_simu_multi2(seed = 1,nvec=nvec, p=p, d= d, 
                                 q=q, qs=qs, rank0=rank0,rho=rho,sigma2_eps=1)
  

```
Fit the MultiCOAP model using the function `MultiCOAP()` in the R package `MultiCOAP`. Users can use `?MultiCOAP` to see the details about this function. For two matrices $\widehat D$ and $D$, we use trace statistic to measure their similarity.  The trace statistic  ranges from 0 to 1,  with higher values indicating better performance. To gauge the estimation accuracy of $\beta_0$, we employ the mean estimation error (Er).
```{r  eval = TRUE}

  XcList <- datList$Xlist; 
  ZList <- datList$Zlist; 
  res <- MultiCOAP(XcList, ZList, q=q, qs=qs,rank_use = rank0,init="MSFRVI")
  #str(res)
  metricMat["MultiCOAP",'Time'] <- res$time.use
  metricMat["MultiCOAP",'A_tr'] <- trace_statistic_fun(res$A, datList$A0)
  metricMat["MultiCOAP",'B_tr'] <- trace_list_fun(res$B, datList$Blist0)
  metricMat["MultiCOAP",'F_tr'] <- trace_list_fun(res$F, datList$Flist)
  metricMat["MultiCOAP",'H_tr'] <- trace_list_fun(res$H, datList$Hlist)
  metricMat["MultiCOAP",'beta_er'] <- normvec(datList$bbeta0-res$bbeta)

```

## Compare with other methods

We compare MultiCOAP with various prominent methods in the literature. They are (1) covariate-augmented Possion factor model (COAP) that disregards the study-specified factors and only estimates the loadings and factors shared across studies, implemented in the R package `COAP`; (2) newly developed multi-study factor regression model (MSFR) with/without log(1+x) normalization for the observed count data. 

(1). First, we implemented covariate-augmented Possion factor model (COAP) and record the metrics that measure the estimation accuracy and computational cost.

```{r  eval = TRUE}
mat2list <-function(z_int, nvec){
  
  zList_int <- list()
  istart <- 1
  for(i in 1:length(nvec)){
    
    zList_int[[i]] <- z_int[istart: sum(nvec[1:i]), ]
    istart <- istart + nvec[i]
  }
  return(zList_int)
}

COAP_run <- function(XcList, ZList, q, rank_use, aList=NULL, ...){
  
  require(COAP)
  nvec <- sapply(XcList, nrow)
  if(is.null(aList)){
    aList <- lapply(nvec, function(n1) rep(1, n1))
  }
  
  Xmat <- Reduce(rbind, XcList); Zmat <- Reduce( rbind, ZList)
  avec <- unlist(aList)
  tic <- proc.time()
  res_coap <- RR_COAP(Xmat, multiFac= avec, Z = Zmat,  q=q, rank_use=rank_use, ...)
  toc <- proc.time()
  res_coap$Flist <- mat2list(res_coap$H, nvec)
  res_coap$time.use <- toc[3] - tic[3]
  return(res_coap)
}

res_coap <- COAP_run(XcList, ZList, q=q,rank_use = rank0)

metricMat["COAP",'Time']  <- res_coap$time.use
metricMat["COAP",'A_tr']  <- trace_statistic_fun(res_coap$B, datList$A0)
metricMat["COAP",'F_tr']  <- trace_list_fun(res_coap$Flist, datList$Flist)
metricMat["COAP",'beta_er']  <- normvec(datList$bbeta0-res_coap$bbeta)
  
```

(2). Then, we implemented multi-study factor regression model (MSFR) with/without log(1+x) normalization for the observed count data.
```{r  eval = TRUE}
MSFR_run <- function(XcList, ZList, q, qs, maxIter=1e4, log.transform=FALSE, truncted=NULL, load.source=FALSE, dir.source=NULL){
  
  require(MSFA)
  require(psych)
  if(!load.source){
    source(paste0(dir.source, "MSFR_main_R_MSFR_V1.R"))
  }
  #fa <- psych::fa
  B_s <- ZList
  if(log.transform){
    X_s <- lapply(XcList, function(x) log(1+x))
  }else{
    X_s <- XcList #
  }
  if(!is.null(truncted)){
    replace_x <- function(x, truncted){
      x[x>truncted] <- truncted
      x[x<-truncted] <- -truncted
      return(x)
    }
    X_s <- lapply(X_s, replace_x, truncted=truncted)
  }
  t1<- proc.time()
  test <- start_msfa(X_s, B_s, 5, k=q, j_s=qs, constraint = "block_lower2", method = "adhoc")
  # EM_beta <- ecm_msfa(X_s, B_s, start=test, trace = FALSE, nIt=maxIter, constraint = "block_lower1")
  EM_beta <- ecm_msfa(X_s, B_s, start=test, trace = FALSE, nIt=maxIter)
  t2<- proc.time()
  EM_beta$Flist <- lapply(EM_beta$E_f, t)
  EM_beta$Hlist <- lapply(EM_beta$E_l, t)
  EM_beta$time.use <- t2[3] - t1[3]
  EM_beta$lambdavec <- sapply(EM_beta$psi_s,mean)
  return(EM_beta)
}
dir.source <- 'D:\\Working\\Research paper\\idea\\MultiPoisFactor\\Rpackage\\MultiCOAP\\simuCodes\\'

res_msfa <- MSFR_run(XcList, ZList, q=q, qs=qs, maxIter=1e3, log.transform=TRUE, dir.source = dir.source)
metricMat["MSFRlog",'Time'] <- res_msfa$time.use
metricMat["MSFRlog",'A_tr'] <- trace_statistic_fun(res_msfa$Phi, datList$A0)
metricMat["MSFRlog",'B_tr']  <- trace_list_fun(res_msfa$Lambda_s, datList$Blist0)
metricMat["MSFRlog",'F_tr'] <- trace_list_fun(res_msfa$Flist, datList$Flist)
metricMat["MSFRlog",'H_tr'] <- trace_list_fun(res_msfa$Hlist, datList$Hlist)
metricMat["MSFRlog",'beta_er'] <- normvec(res_msfa$beta-datList$bbeta0)
  
res_msfa2 <- MSFR_run(XcList, ZList, q=q, qs=qs,truncted=500, maxIter=1e3, log.transform=FALSE, dir.source = dir.source)
metricMat["MSFR",'Time'] <- res_msfa2$time.use
metricMat["MSFR",'A_tr'] <- trace_statistic_fun(res_msfa2$Phi, datList$A0)
metricMat["MSFR",'B_tr']  <- trace_list_fun(res_msfa2$Lambda_s, datList$Blist0)
metricMat["MSFR",'F_tr'] <- trace_list_fun(res_msfa2$Flist, datList$Flist)
metricMat["MSFR",'H_tr'] <- trace_list_fun(res_msfa2$Hlist, datList$Hlist)
metricMat["MSFR",'beta_er'] <- normvec(res_msfa2$beta-datList$bbeta0)
    
```

## Visualize the comparison of performance

Next, we summarized the metrics for MultiCOAP and other compared methods in a data.frame object.
```{r  eval = TRUE}


dat_metric <- data.frame(metricMat)
dat_metric$Method <- factor(row.names(dat_metric), levels=row.names(dat_metric))
```

Plot the results for MultiCOAP and other methods, which suggests that MultiCOAP achieves better estimation accuracy for the quantiites of interest.
```{r  eval = TRUE, fig.width=13, fig.height=5}
library(cowplot)
library(ggplot2)
p1 <- ggplot(data=subset(dat_metric, !is.na(A_tr)), aes(x= Method, y=A_tr, fill=Method)) + geom_bar(stat="identity") + xlab(NULL) + scale_x_discrete(breaks=NULL) + theme_bw(base_size = 16)
p2 <- ggplot(data=subset(dat_metric, !is.na(F_tr)), aes(x= Method, y=F_tr, fill=Method)) + geom_bar(stat="identity") + xlab(NULL) + scale_x_discrete(breaks=NULL)+ theme_bw(base_size = 16)
p3 <- ggplot(data=subset(dat_metric, !is.na(B_tr)), aes(x= Method, y=B_tr, fill=Method)) + geom_bar(stat="identity") + xlab(NULL) + scale_x_discrete(breaks=NULL) + theme_bw(base_size = 16)
p4 <- ggplot(data=subset(dat_metric, !is.na(H_tr)), aes(x= Method, y=H_tr, fill=Method)) + geom_bar(stat="identity") + xlab(NULL) + scale_x_discrete(breaks=NULL)+ theme_bw(base_size = 16)
p5 <- ggplot(data=subset(dat_metric, !is.na(beta_er)), aes(x= Method, y=beta_er, fill=Method)) + geom_bar(stat="identity") + xlab(NULL) + scale_x_discrete(breaks=NULL)+ theme_bw(base_size = 16)
p6 <- ggplot(data=subset(dat_metric, !is.na(Time)), aes(x= Method, y=Time, fill=Method)) + geom_bar(stat="identity") + xlab(NULL) + scale_x_discrete(breaks=NULL)+ theme_bw(base_size = 16)
plot_grid(p1,p2,p3, p4, p5, p6, nrow=2, ncol=3)
```


## Select the parameters

We applied the proposed 'CUP' method to select the number of factors. The results showed that  the CUP method has the potential to identify the true values.
```{r  eval = TRUE}
selectQ <- function(res, tune.coef.rank=FALSE,  threshold=c(1e-1, 1e-2, 1e-2), 
                    method=c("var.prop", "SVR"), upper.var.prop=0.95){
  
  method <- match.arg(method)
  qlist <- list()
  
  d_svdA <- svd(res$A)$d
  # 
  if(method=='SVR'){
    d_svdA <- d_svdA[d_svdA>threshold[1]]
    qq <- length(d_svdA)
    if(qq>1){
      rsigs <- d_svdA[-qq]/d_svdA[-1]
      qlist$q <-  which.max(rsigs)
    }else{
      qlist$q <- 1
    }
  }else if(method=='var.prop'){
    prop.vec <- cumsum(d_svdA^2)/sum(d_svdA^2)
    qlist$q <- which(prop.vec> upper.var.prop)[1]
  }
  
  n_qs <- length(res$B)
  qvec <- rep(NA, n_qs)
  names(qvec) <- paste0("qs", 1:n_qs)
  for(i in 1:n_qs){
    # i <-1 
    d_svdB1 <- svd(res$B[[i]])$d
    # 
    if(method=='SVR'){
      d_svdB1 <- d_svdB1[d_svdB1>threshold[2]]
      qq1 <- length(d_svdB1)
      if(qq1>1){
        rsigB1 <- d_svdB1[-qq1]/d_svdB1[-1]
        qvec[i] <-  which.max(rsigB1)
      }else{
        qvec[i] <-  1
      }
    }else if(method=='var.prop'){
      prop.vec <- cumsum(d_svdB1^2)/sum(d_svdB1^2)
      qvec[i] <- which(prop.vec> upper.var.prop)[1]
    }
  }
  qlist$qs <- qvec
  return(qlist)
}

XcList <- datList$Xlist; 
ZList <- datList$Zlist; 
res <- MultiCOAP(XcList, ZList, q=q, qs=qs,rank_use = rank0,init="MSFRVI", verbose = FALSE)
hq.list <- selectQ(res, method='var.prop')
message("hq = ", hq.list$q, " VS true q = ", q)
message("hqs.vec = ", paste(hq.list$qs, collapse =", "), " VS true qs.vec = ", paste(qs, collapse =", "))
```


<details>
<summary>**Session Info**</summary>
```{r}
sessionInfo()
```
</details>